Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children

### 2023-10-20 manuscript/main.tex
Guilherme comments
short title and rearrange post-text

\subsection*{Teaser}

@@

We investigated the co-development of the brain and microbiome in
381 healthy and neurotypically-developing children (172 female) between 40 days and 10 years of age
(table 1, fig. 1A, fig. S1)
using a variety of orthogonal microbial and neurocognitive measures.
These included shotgun metagenomic sequencing,
age-appropriate cognitive and behavioral assessments
using full-scale composite scores from the MSEL, WPPSI-4 and WISC-V,
and neuroimaging measures of cortical and sub-cortical morphology. 
Different assessment instruments were used depending on the child's age
due to the unique age-range and psychometric properties measured by each
(i.e., the MSEL for children under 4 years of age, The WPPSI-3 for 3-6 year-olds,
and the WISC-V for children older than 6), but were normalized to a common scale (fig. 1B),
As expected, the greatest differences in microbial taxa were observed across
child age (fig. 1B, E; PCoA axis 1), with older children primarily
stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
high-abundance \emph{Prevotella copri} (fig. S2).
Overall subject-to-subject variation in gut microbial genes
and brain volume profiles was similarly driven largely by subject age (fig. 1C, F).

@@

As with LMs, RF models for children under 6 months old
performed poorly (mean test-set correlation -0.13, mean RMSE 13.01),
but RFs were consistently able to learn the relationship between taxa
and cognitive function scores in children over 18 months of age (mean
test-set correlation 0.429, mean RMSE 18.25).
Species that were important in RF overlapped with those that were
significant when testing the relationship with LMs (fig.
3A-B, table S1A-B); Both analysis identified 
\textit{F. prausnitzii}, % F. prausz was mentioned before, ~ line 395
\textit{E. eligens}, % E. eligens was mentioned before, ~ line 395
\textit{Parasutterella excrementihominis}, and % Not mentioned before
\textit{A. obesi} as important % A. obesi was mentioned before, ~ lines 120, 390
(q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs).
Other taxa that were significant in LMs were not identified by RF models
including \textit{Eggerthella} sp. CAG 298,
and \textit{B. intestinihominis}. \textit{E. ramulus}
was just outside our cutoff (64\% cumulative importance). % Flag - 64%?
RF models also identified an additional 48 species
that were important for discriminating cognitive function scores
(table 3, fig. 3A-B, E). Subject age was consistently ranked highly in
feature importance, which could indicate that decision branches based on
microbial taxa have increased purity when considering the subject's age
or that age itself is a useful predictor.

@@

As is made apparent by the potentially opposing roles of 
\textit{B. fragilis} and \textit{B. ovatus},
individual species from the same genus may play different roles.
As such, the use of shotgun metagenomic sequencing
may be crucial to identifying microbial influences on the brain,
since it enables species-level resolution of microbial taxa.
A previous study of
cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing
and showed that genera from the Lachnospiraceae family as well as
unclassified Clostridiales (now Eubacteriales) were associated with
higher scores on the Ages and Stages Questionnaire
\cite{sordilloAssociationInfantGut2019}.
However, each of these clades encompass dozens of genera with
diverse functions, each of which may have different effects. Indeed,
several of the taxa that were positively associated with cognitive
function in this study, including \emph{B. wexlerae}, \emph{D.
longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
Clostridiales, as is \emph{R. gnavus}, which we found was negatively
associated with cognitive function (fig. 2A-B). This kind of
species-level resolution is typically not possible with amplicon
sequencing.

@@

\subsubsection*{Data and materials availability}

@@

\subsubsection*{Funding}
This research was funded by NIH UG3 OD023313 and Wellcome LEAP 1kD.

\subsubsection*{Competing Interests}
The authors declare that they have no competing interests.

\subsubsection*{Author contributions}
% https://credit.niso.org/

\subsection*{Figures}

>>>>

\subsubsection*{Short title}

Gut microbes are linked to child neurocognition

\subsubsection*{Teaser}

@@

We investigated the co-development of the brain and microbiome in
381 healthy and neurotypically-developing children (172 female) between 40 days and 10 years of age
(table 1, fig. 1A, fig. S1)
using a variety of orthogonal microbial and neurocognitive measures.
These included shotgun metagenomic sequencing,
age-appropriate cognitive and behavioral assessments
using full-scale composite scores from the MSEL, WPPSI-4 and WISC-V,
and neuroimaging measures of cortical and sub-cortical morphology. 
Different assessment instruments were used depending on the child's age
due to the unique age-range and psychometric properties measured by each
(i.e., the MSEL for children under 4 years of age, The WPPSI-3 for 3-6 year-olds,
and the WISC-V for children older than 6), but were normalized to a common scale (fig. 1D),
As expected, the greatest differences in microbial taxa were observed across
child age (fig. 1B, E; PCoA axis 1), with older children primarily
stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
high-abundance \emph{Prevotella copri} (fig. S2).
Overall subject-to-subject variation in gut microbial genes
and brain volume profiles was similarly driven largely by subject age (fig. 1C, F).

@@

As with LMs, RF models for children under 6 months old
performed poorly (mean test-set correlation -0.13, mean RMSE 13.01),
but RFs were consistently able to learn the relationship between taxa
and cognitive function scores in children over 18 months of age (mean
test-set correlation 0.429, mean RMSE 18.25).
Species that were important in RF overlapped with those that were
significant when testing the relationship with LMs (fig.
3A, table S1); Both analysis identified 
\textit{F. prausnitzii}, % F. prausz was mentioned before, ~ line 395
\textit{E. eligens}, % E. eligens was mentioned before, ~ line 395
\textit{Parasutterella excrementihominis}, and % Not mentioned before
\textit{A. obesi} as important % A. obesi was mentioned before, ~ lines 120, 390
(q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs).
Other taxa that were significant in LMs were not identified by RF models
including \textit{Eggerthella} sp. CAG 298,
and \textit{B. intestinihominis}. \textit{E. ramulus}
was just outside our cutoff (64\% cumulative importance). % Flag - 64%?
RF models also identified an additional 48 species
that were important for discriminating cognitive function scores
(table 3, fig. 3B, E). Subject age was consistently ranked highly in
feature importance, which could indicate that decision branches based on
microbial taxa have increased purity when considering the subject's age
or that age itself is a useful predictor.

@@

As is made apparent by the potentially opposing roles of 
\textit{B. fragilis} and \textit{B. ovatus},
individual species from the same genus may play different roles.
As such, the use of shotgun metagenomic sequencing
may be crucial to identifying microbial influences on the brain,
since it enables species-level resolution of microbial taxa.
A previous study of
cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing
and showed that genera from the Lachnospiraceae family as well as
unclassified Clostridiales (now Eubacteriales) were associated with
higher scores on the Ages and Stages Questionnaire
\cite{sordilloAssociationInfantGut2019}.
However, each of these clades encompass dozens of genera with
diverse functions, each of which may have different effects. Indeed,
several of the taxa that were positively associated with cognitive
function in this study, including \emph{B. wexlerae}, \emph{D.
longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
Clostridiales, as is \emph{R. gnavus}, which we found was negatively
associated with cognitive function (fig. 3C, F). This kind of
species-level resolution is typically not possible with amplicon
sequencing.

@@

\section*{Funding}
This research was funded by NIH UG3 OD023313 and Wellcome LEAP 1kD.

\section*{Author contributions}
% https://credit.niso.org/

\section*{Competing Interests}
The authors declare that they have no competing interests.
\section*{Figures}

\section*{Data and materials availability}

@@
